Software Citation Principles

### 2016-03-08 sc-principles/software-citation-principles.tex
edit in motivation of why software citation is not just a style issue
editing in list of motivations in section 2

As the process of science has become increasingly digital, scientific outputs and products have
grown beyond simple papers and books to include software, data, and other electronic
components.  Scientific knowledge is embedded in these components.  And papers and books
themselves are also becoming increasingly digital, allowing them to become executable
and reproducible.  As we move towards this future where science is performed in and recorded
as a variety of linked digital products, the characteristics and properties that developed for
books and papers need to be applied to all digital products and possibly adjusted.  Here, we are concerned specifically
with the citation of software products.  This is more general than just textual citations of software in a paper, but more general identification of software within the research process. \katznote{other comments on this last sentence are welcome - is this a good enough description of what we mean by software citation?}

@@

There are many reasons this lack of both software citations in general and standard practices
for software citation are of concern:
\begin{itemize}
\item Understanding Science: Software is a product of research, and by not citing it, we leave holes in the record of scientific progress.
\item Academic Credit: Academic researchers at all levels, including students, postdocs, faculty, and staff, should be
credited for the software products they develop and contribute to, particularly when those products enable or further research done by others.
\item Finding Software - citations enable software to be discovered \LJHnote{and reused}. \katznote{dan to flesh out more} \ssnote{is it worth framing as software(cite \& find)->researcher(credit)->science (reuse) areas?}
\item Reproducibility - citations enable specific software that was
used to be identified, though they are not sufficient to make reproducibility happen \katznote{better language welcome}
\end{itemize}

\acmnote{would like to add info on the problem we're trying to solve in academia to motivation} 
\katznote{added list above - others can add more items if appropriate.}

>>>>

As the process of science has become increasingly digital, scientific outputs and products have
grown beyond simple papers and books to include software, data, and other electronic
components.  Scientific knowledge is embedded in these components.  And papers and books
themselves are also becoming increasingly digital, allowing them to become executable
and reproducible.  As we move towards this future where science is performed in and recorded
as a variety of linked digital products, the characteristics and properties that developed for
books and papers need to be applied to all digital products and possibly adjusted.  Here, we are concerned specifically
with the citation of software products.  The challenge is not just the textual citation of software in a paper, but the more general identification of software within the research process. \katznote{other comments on this last sentence are welcome - is this a good enough description of what we mean by software citation?}

@@

There are many reasons this lack of both software citations in general and standard practices
for software citation are of concern: \katznote{Any more items to be added to list?}

\begin{itemize}
\item Understanding Science: Software is a product of research, and by not citing it, we leave holes in the record of scientific progress.
\item Academic Credit: Academic researchers at all levels, including students, postdocs, faculty, and staff, should be
credited for the software products they develop and contribute to, particularly when those products enable or further research done by others.
\item Finding Software: Citations enable the specific software used in a research product to be discovered. Additional researchers can then use the same software for different purposes, leading
to credit for those responsible for the software.
\item Reproducibility: Citation of specific software used is necessary for reproducibility, but is not sufficient.  Additional information such as configurations and platform issues are also needed.
\end{itemize}
